---
title: "p8105_hw5_sm4956"
author: Shivangi Deepak Mewada
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = TRUE)
```
Loading the libraries required for the midterm, configuring the code chunk dimensions, and creating themes and display for the plots for knit.


## Problem 0
* created repository and R project for HW5, created rmd file and rending to GitHub. 
* created a sub-directory/ data folder for the data set files to be used for this HW

```{r load libraries}
library(tidyverse)
library(readxl)
library(dplyr)
library(ggridges)
library(tibble)
library(broom)
library(purrr)
options(tibble.print_min = 5)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

Importing data in individual spreadsheets from data subdirectory. The dataframe was created that includes list of all files in the directory. map used to import data using `read_csv` function; and unnest used.
```{r}
full_df = 
  tibble(
    files = list.files("data/zip_data/"),
    path = str_c("data/zip_data/", files)
  ) %>% 
  mutate(data = map(path, read_csv)) %>% 
  unnest(cols = c(data))
```

Tidying the data, manipulating file names to include control arm and subject ID, weekly observations tidied, with other data wrangling.
```{r}
tidy_df = 
  full_df %>% 
  mutate(
    files = str_replace(files, ".csv", ""),
    group = str_sub(files, 1, 3)) %>% 
  pivot_longer(
    week_1:week_8,
    names_to = "week",
    values_to = "outcome",
    names_prefix = "week_") %>% 
  mutate(week = as.numeric(week)) %>% 
  select(group, subj = files, week, outcome)
```

Spaghetti plot to show observations on each subject over time, and noting the trend
```{r}
tidy_df %>% 
  ggplot(aes(x = week, y = outcome, group = subj, color = group)) + 
  geom_point() + 
  geom_path() + 
  facet_grid(~group)
```

This plot suggests high within-subject correlation -- subjects who start above average end up above average, and those that start below average end up below average. Subjects in the control group generally don't change over time, but those in the experiment group increase their outcome in a roughly linear way. 
